### 2016-11-05 introduction.tex
saturday more shit

Essentially, \cite{mozer1989using} observed that neural networks during training do \textit{not} distribute the learning representation evenly or equitably across hidden units. What actually happens is that a few, elite neurons learn an approximation of the input-output function, and the remaining units must learn a complex interdependence function which cancels out their respective influence on the network output. Furthermore, assuming enough units exist to learn the function in question, increasing the number of parameters does not increase the richness or robustness of the learned approximation, but rather simply increases the number of noisy parameters to be canceled during training. This is evinced by the fact that in many cases, multiple neurons can be removed from a network with no re-training and with negligible impact on the quality of the output approximation. In other words, there are few bipartisan units in a trained network. A unit is typically either part of the input-output function approximation, or it is part of an elaborate noise cancellation task force. Assuming this is the case, most of the compute-time spent training a neural network is likely occupied by this arguably wasteful procedure of silencing superfluous parameters, and pruning can be viewed as a necessary treatment procedure to ``trim the fat.''

We observed copious evidence of this phenomenon in our experiments, and this is the motivation behind our decision to evaluate the pruning algorithms in this study on the simple criteria of their ability to trim neurons \textit{without} any re-training. If we were to employ re-training as part of our evaluation criteria, we would arguably \textit{not} be evaluating the quality of our algorithm's pruning decisions per se but rather the ability of back-propagation trained networks to recover from faults caused by non-ideal pruning decisions, as suggested by the conclusions of \cite{segee1991fault} and \cite{mozer1989skeletonization}. Moreover, as \cite{fahlman1989cascade} discuss, due to the ``herd effect'' and ``moving target'' phenomena in back-propagation learning, the remaining units in a network will simply shift course to account for whatever error signal is re-introduced as a result of a bad pruning decision or network fault. So long as there are enough critical parameters to learn the function in question, a network can typically recover with additional training. This limits the conclusions we can draw about the quality of our pruning criteria when we employ re-training.

@@

If we wanted to continually shrink a network to its absolute minimum possible size, we might accomplish this using any number of off-the-shelf pruning algorithms, such as Skeletonization (\cite{mozer1989skeletonization}), Optimal Brain Damage (\cite{lecun1989optimal}), or later variants such as Optimal Brain Surgeon (\cite{hassibi1993second}). In fact, we borrow much of our inspiration from these algorithms, with one major variation: Instead of pruning individual weights, we prune entire neurons, thereby eliminating all of their incoming and outgoing weight parameters in one go, resulting in more memory saved, faster.

Scoring and ranking individual weight parameters in a large network is computationally expensive, and generally speaking the removal of a single weight from a large network is a drop in the bucket in terms of reducing a network's core memory footprint.  We argue that pruning neurons instead of weights is more efficient computationally as well as practically in terms of quickly reaching a target reduction in memory size. Our approach also attacks the angle of giving downstream applications a realistic expectation of the minimal increase in error resulting from the removal of a specified percentage of neurons. Such trade-offs are unavoidable, but performance impacts can be limited if a principled approach is used to find candidate neurons for removal.

Too many free parameters in a Neural Network lead to overfitting. Regardless of the number of weights used in a given network, as \cite{segee1991fault} assert, the representation of a learned function approximation is almost never evenly distributed over the hidden units, and the removal of any single hidden unit at random can actually result in a total network fault. \cite{mozer1989using} suggest that only a subset of the hidden units in a neural network actually latch on to the invariant or generalizing properties of the training inputs, and the rest learn to either mutually cancel each other out or begin over-fitting to the noise in the data. We leverage this idea in the current work to rank all neurons in pre-trained networks based on their effective contributions to the overall performance. We then remove the unnecessary neurons to reduce the network's footprint. Through our experiments we not only concretely validate the theory put forth by \cite{mozer1989using} but we also successfully build on  it to prune networks to 40 to 60 \% of their original size without any major loss in performance.

>>>>

Essentially, \cite{mozer1989using} observed that neural networks during training do \textit{not} distribute the learning representation evenly or equitably across hidden units. What actually happens is that a few, elite neurons learn an approximation of the input-output function, and the remaining units must learn a complex interdependence function which cancels out their respective influence on the network output. Furthermore, assuming enough units exist to learn the function in question, increasing the number of parameters does not increase the richness or robustness of the learned approximation, but rather simply increases the likelihood of overfitting and the number of noisy parameters to be canceled during training. This is evinced by the fact that in many cases, multiple neurons can be removed from a network with no re-training and with negligible impact on the quality of the output approximation. In other words, there are few bipartisan units in a trained network. A unit is typically either part of the (possibly overfit) input-output function approximation, or it is part of an elaborate noise cancellation task force. Assuming this is the case, most of the compute-time spent training a neural network is likely occupied by this arguably wasteful procedure of silencing superfluous parameters, and pruning can be viewed as a necessary treatment procedure to ``trim the fat.''

We observed copious evidence of this phenomenon in our experiments, and this is the motivation behind our decision to evaluate the pruning algorithms in this study on the simple criteria of their ability to trim neurons \textit{without} any re-training. If we were to employ re-training as part of our evaluation criteria, we would arguably \textit{not} be evaluating the quality of our algorithm's pruning decisions per se but rather the ability of back-propagation trained networks to recover from faults caused by non-ideal pruning decisions, as suggested by the conclusions of \cite{segee1991fault} and \cite{mozer1989skeletonization}. Moreover, as \cite{fahlman1989cascade} discuss, due to the ``herd effect'' and ``moving target'' phenomena in back-propagation learning, the remaining units in a network will simply shift course to account for whatever error signal is re-introduced as a result of a bad pruning decision or network fault. So long as there are enough critical parameters to learn the function in question, a network can typically recover faults with additional training. This limits the conclusions we can draw about the quality of our pruning criteria when we employ re-training.

@@

If we wanted to continually shrink a neural network down to minimum size, the most straightforward brute-force way to do it is to individually switch each element off and measure the increase in total error on the training set. We then pick the element which has the least impact on the total error, and remove it. Rinse and repeat. This is extremely computationally expensive, given a reasonably large neural network and training set. Alternatively, we might accomplish this using any number of much faster off-the-shelf pruning algorithms, such as Skeletonization (\cite{mozer1989skeletonization}), Optimal Brain Damage (\cite{lecun1989optimal}), or later variants such as Optimal Brain Surgeon (\cite{hassibi1993second}). In fact, we borrow much of our inspiration from these algorithms, with one major variation: Instead of pruning individual weights, we prune entire neurons, thereby eliminating all of their incoming and outgoing weight parameters in one go, resulting in more memory saved, faster.

The algorithm developed for this paper is targeted at reducing the total number of neurons in a trained network, which is one way of reducing its computational memory footprint. This is often a desirable criteria to minimize in the case of resource-constrained or embedded devices, and also allows us to probe the limitations of pruning down to the very last essential network elements. In terms of generalization as well, we can measure the error of the network on the test set as each element is sequentially removed from the network. With an oracle pruning algorithm, what we expect to observe is that the output of the network remains stable as the first few superfluous neurons are removed, and as we start to bite into the more crucial members of the function approximation, the error should start to rise dramatically. In this paper, the brute-force approach described at the beginning of this section serves as a proxy for an oracle pruning algorithm.

One reason to choose to rank and prune individual neurons as opposed to weights is that there are far fewer elements to consider. Furthermore, the removal of a single weight from a large network is a drop in the bucket in terms of reducing a network's core memory footprint. If we want to reduce the \textit{size} of a network as efficiently as possible, we argue that pruning neurons instead of weights is more efficient computationally as well as practically in terms of quickly reaching a hypothetical target reduction in memory consumption. This approach also offers downstream applications a realistic expectation of the minimal increase in error resulting from the removal of a specified percentage of neurons. Such trade-offs are unavoidable, but performance impacts can be limited if a principled approach is used to find the best candidate neurons for removal.

It is well known that too many free parameters in a neural network can lead to overfitting. Regardless of the number of weights used in a given network, as \cite{segee1991fault} assert, the representation of a learned function approximation is almost never evenly distributed over the hidden units, and thus the removal of any single hidden unit at random can actually result in a network fault. \cite{mozer1989using} argue that only a subset of the hidden units in a neural network actually latch on to the invariant or generalizing properties of the training inputs, and the rest learn to either mutually cancel each other's influence or begin over-fitting to the noise in the data. We leverage this idea in the current work to rank all neurons in pre-trained networks based on their effective contributions to the overall performance. We then remove the unnecessary neurons to reduce the network's footprint. Through our experiments we not only concretely validate the theory put forth by \cite{mozer1989using} but we also successfully build on  it to prune networks to 40 to 60 \% of their original size without any major loss in performance.

### 2016-11-05 methodology.tex
saturday more shit

\section{Pruning Neurons to Shrink Neural Networks}\label{sec2}
As discussed in Section \ref{sec1} our aim here is to leverage the non-uniform distribution of learning representation in pre-trained neural networks to eliminate non-contributing neurons in their entirety, without focusing on individual weight parameters. Taking this approach enables us to remove all the weights (incoming and outgoing) associated with a non-contributing neuron in one go. We would like to note here that in an ideal scenario, based on the neuron interdependency theory put forward by \cite{mozer1989skeletonization}, one would evaluate all possible combinations of neurons to remove (one at a time, two at a time, three at a time and so forth) to find the optimal subset of neurons to keep. This however is not feasible practically, which is why we will only focus on removing one neuron at a time and explore "greedy" algorithms to do this in an efficient manner.

>>>>

\section{Pruning Neurons to Shrink Neural Networks}\label{sec2}
As discussed in Section \ref{sec1} our aim here is to leverage the non-uniform distribution of the learning representation in pre-trained neural networks to eliminate excess neurons in their entirety, without focusing on individual weight parameters. Taking this approach enables us to remove all the weights (incoming and outgoing) associated with a non-contributing neuron in one go. We would like to note here that in an ideal scenario, based on the neuron interdependency theory put forward by \cite{mozer1989skeletonization}, one would evaluate all possible combinations of neurons to remove (one at a time, two at a time, three at a time and so forth) to find the optimal subset of neurons to keep. This however is not feasible practically, which is why we will only focus on removing one neuron at a time and explore "greedy" algorithms to do this in an efficient manner.

### 2016-11-05 results.tex
saturday more shit

\section{Experimental Results}
For all the results presented in this section, the MNIST database of Handwritten Digits by \cite{lecun-mnisthandwrittendigit-2010} was used. It is worth noting here that since only a subset of the entire dataset was used in order to speed up experiments, the starting test accuracies reported here might differ from those reported by LeCun. This does not impact the presented results, since it's the change in the error and accuracy of the network post pruning which is of significance here and not the starting accuracy itself.

\subsection{Pruning a 1-Layer Network}
The network architecture in this case consisted of 1 layer, 100 neurons, 10 outputs, logistic sigmoid activations, and a starting test accuracy of 0.998.
\subsubsection{Single Overall Ranking Algorithm}
We first present the results for a single-layer neural network in Figure \ref{fig:mnist-single-ranking-single-layer}, using the Single Overall algorithm (Algorithm \ref{algo1}) as proposed in Section \ref{sec2}. (We again note that this algorithm is intentionally naive and is used for comparison only. Its performance should be expected to be poor.) After training, each neuron is assigned its permanent ranking based on the three criteria discussed previously: A brute force ``ground truth'' ranking, and two approximations of this ranking using first and second order Taylor estimations of the change in network output error resulting from the removal of each neuron.

>>>>

\section{Experimental Results}
For all the results presented in this section, the MNIST database of Handwritten Digits by \cite{lecun-mnisthandwrittendigit-2010} was used. It is worth noting that due to the time taken by the brute force algorithm we rather used a 5000 image subset of the MNIST database in which we have normalized the pixel values between 0 and 1.0, and compressed the image sizes to 20x20 images rather than 28x28, so the starting test accuracy reported here appears higher than those reported by LeCun et al. We do not believe that this affects the interpretation of the presented results because the basic learning problem does not change with a larger dataset or input dimension.

\subsection{Pruning a 1-Layer Network}
The network architecture in this case consisted of 1 layer, 100 neurons, 10 outputs, logistic sigmoid activations, and a starting test accuracy of 0.998.

\subsubsection{Single Overall Ranking Algorithm}
We first present the results for a single-layer neural network in Figure \ref{fig:mnist-single-ranking-single-layer}, using the Single Overall algorithm (Algorithm \ref{algo1}) as proposed in Section \ref{sec2}. (We again note that this algorithm is intentionally naive and is used for comparison only. Its performance should be expected to be poor.) After training, each neuron is assigned its permanent ranking based on the three criteria discussed previously: A brute force ``ground truth'' ranking, and two approximations of this ranking using first and second order Taylor estimations of the change in network output error resulting from the removal of each neuron.
